Stack Safety as a Security Property

### 2020-07-01 main.tex
Example in trace confidentiality.

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). \leo{There needs to be some informal description
  of the wbcf part here. Sean?}

@@

In this section we will describe stack safety as a formal property. We
begin with introducing a running example that will help keep things
concrete throughout this section, continue with our model of
\leo{compiler-provided?} knowledge about the program structure, and
then dive in to formal definitions local state encapsulation and
well-bracketed control flow.

\begin{figure}[b]
  \begin{centering}
    \begin{minipage}{0.5\textwidth}
\begin{verbatim}
f{
  int x, y;
  ...
  y = g(0);
  ...
}

g(x){
  return (x + h(17));
}

h(x){
  return (x + 1);
}
\end{verbatim}
\end{minipage}
\end{centering}
\caption{High-level Program with Simple Nested Calls}
\label{fig:running-program}
\end{figure}

@@

Our running example consists of three functions $f$, $g$, and $h$ that
perform a nested sequence of calls. A high level C-like view of the
excerpt program can be found in Figure~\ref{fig:running-program}.
Suppose that the starting state for our running example is right
before $f$'s call to $g$. A rough example layout of the memory at that
point can be found in the top of Figure~\ref{fig:running-trace-b}:
\begin{itemize}
\item The low part of the memory (0-1000) is dedicated to instructions.
\item The stack starts at location 1000.
\item The stack of $f$ starts at 1040.
\item The location of $x$ is 1040 and stores $42$.
\item The location of $y$ is 1044 and stores $0$.
\end{itemize}
Directly below the layout, we summarize the parts of the register file
that are relevant to the example.

The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Unoptimized, etc.}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
\newcommand*{\rsp}{\mathit{sp}}
\newcommand*{\rra}{\mathit{ra}}

\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}

\begin{figure}

%Starting Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ? 
\end{array}
\]
\vspace*{0.2em}
\[
  \begin{array}{r|c|c|c|c}
    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
    \hline
    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}
    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
  \end{array}
  \]
~ \\
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ? 
\end{array}
\]
% G - Body - 1
\vspace*{0.2em}
\[
  \begin{array}{r|c|c|c|c}
    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}
    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}
    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
  \end{array}
  \]
  ~ \\
% At Call to H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{0}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ? 
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
\]
% At Middle of H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{52em}{1.8em}{$\downarrow$}
\MemoryLabel{51.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{1}{cyan}%
\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{0}
\MemoryLabel{-14.0em}{0.6em}{168}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 212 & r_0 : 0 & \rra : 168 & \rsp : 1072 & r_4 : 18 & r_5 : ? 
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
  \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
  \end{array}
\]
\label{fig:running-trace-a}
\end{figure}

\begin{figure}
% After Return from H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ? 
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
\end{array}
\]
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
\end{array}
\]
\vspace*{0.2em}  
% Ending Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
\end{array}
\]
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}

%\end{minipage}
%\begin{minipage}{0.7\textwidth}
%\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
%\newcommand*{\clabel}[1]{#1: & & &\\}
%\[
%\begin{array}{llll}
%  \clabel{TOP}  
%  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
%  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
%  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
%  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
%  & & & \\
%  & & & \\
%  \clabel{F}
%  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
%  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
%  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
%  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
%  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
%  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
%  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
%  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
%  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
%  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
%  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
%  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
%  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
%  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
%  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
%  & & & \\
%  & & & \\  
%  \clabel{G}
%  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
%  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
%  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
%  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
%  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
%  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
%  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
%  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
%\end{array}
%\]
%\end{minipage}
%\label{fig:running-program}
%\end{figure}

@@

Formally a code map is a partial function from addresses to a function
identifier drawn from the set \(\FUNIDS\).
  %
\[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
%
In our running example, instructions 100-124 correspond to $f$,
instructions 140-196 correspond to $g$, and instructions 200-228
correspond to $h$.

A call map is a partial function from addresses to a natural number,
the number of arguments and return values of the call.
  %
\[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
%
In our running example, the callmap would contain (at least) two
entries: the word 112 corresponding to the first $\jal$
instruction to $g$, which would map to $2$ as the call has one
argument; and the word 164 also mapping to $2$ corresponding
to the call to $h$.

@@

A return map is simply a predicate on addresses and may be considered
to identify the final jump of the return sequence. An entry map is
likewise a predicate on addresses.
  %
\[\retmap \in \RETMAPS \subseteq \WORDS\]
\[\entmap \in \ENTMAPS \subseteq \WORDS\]
%
In our running example, the return map identifies the $\jalr$
instructions 196 and 224, while the entry map identifies
instructions 140 and 200.

@@

    \definition{Observable Confidentiality}

      Observable confidentiality is concerned with whether high confidentiality
      data influences observable both behavior during and after a call. As with
      eager confidentiality, we will compare the behavior of the trace induced
      by a state to another induced by its variant. But if the original returns,
      before continuing, we must reset the variant to conform with the original
      state.

>>>>

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establshes safety between
functions: all code is owned by a function, and all jumps between functions
are either calls that enter at a marked entry point or returns that enter
just after their associated call in the standard nested call structure.

%\leo{There needs to be some informal description
%  of the wbcf part here. Sean?}

@@

In this section we will describe stack safety as a formal property. We
begin with our model of knowledge about the program structure\leo{I
  hate this}, and then dive in to local state encapsulation and
well-bracketed control flow.

@@

\leo{TODO: Figure out presentation}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
%\begin{lstlisting}
%START:
%  addi r2 r2 12 ; allocate 3 words in the stack
%  sw r2 42 -8; store 42 in a private location
%  sw r2 0 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 F ; call F and store return address
% 
%F:
%  addi r2 r2 4 ; allocate for ra 
%  sw r2 r1 -4 ; store the return address to sp-4
%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address
% 
%G:  
%  addi r2 r2 4 ; allocate for ra 
%  sw r2 r1 -4 ; store the return address to sp-4
%  
%\end{lstlisting}

Starting State:

Instruction memory in locations 0 - ..
PC at 100.
Stack memory starting at 1000.
Assume we're in a function. 
Stack Pointer at 1040.
1040: 42.

\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-18.7em}{0.6em}{42}
\MemoryLabel{-16.4em}{0.6em}{0}
\\
\end{center}

Program:

\begin{verbatim}
top{
  return f(0);
}

f(x){
  return (x + g(17));
}

g(x){
  return (x + 1);
}
\end{verbatim}

\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
\newcommand*{\clabel}[1]{#1: & & &\\}
\newcommand*{\rsp}{sp}
\newcommand*{\rra}{ra}
\[
\begin{array}{llll}
  \clabel{TOP}  
  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
  & & & \\
  & & & \\
  \clabel{F}
  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
  & & & \\
  & & & \\  
  \clabel{G}
  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}

%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address

\end{array}
\]

@@

Formally a code map is a partial function from addresses to a function
identifier drawn from the set \(\FUNIDS\).
  %
\[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]

\leo{Visible from assembly}

A call map is a partial function from addresses to a natural number,
the number of arguments of the call.
  %
\[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]

\leo{The two JAL instructions map to 2 (arg + return)}

@@

A return map is simply a predicate on addresses and may be considered
to identify the final jump of the return sequence. An entry map is
likewise a predicate on addresses.
  %
\[\retmap \in \RETMAPS \subseteq \WORDS\]
\[\entmap \in \ENTMAPS \subseteq \WORDS\]

\leo{retmap is the jal + 1}
\leo{entrymap is the ones after the labels}

@@

    \definition{Observable Trace Confidentiality}

      Trace confidentiality is a property of two traces whose initial machine
      states are variations, with two sub-properties. First, the behavioral
      condition -- that their observable behaviors prior to the end of the
      trace are equivalent -- reflects the intuition that secret values should
      not influence visible behavior. Second, we have the end condition that
      their final states (if any) are equal up to the original variation. The
      end condition over-approximates the extension of the behavioral condition
      to future execution.

      We introduce {\em observable trace confidentiality}, which extends the
      behavioral condition to apply to execution from the final state on.
      This requires restoring varied components to their original values,
      since after return they will once again be accessible legitimately.

\begin{figure}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\
  ~\\
  ~\\
  \begin{verbatim}
    // read HC data and write based on it
    lw r1 r0 40
    bnz r1 2
    sw 0 r0 42
    jmp r2 // return
    sw 1 r0 42
    jmp r2 // return
  \end{verbatim}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{red}
      \memory{1}{black}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{yellow}%
      \memory{1}{white}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\

  \caption{Failing Trace Confidentiality}
  \label{fig:lazyex}
\end{figure}

      In Figure \ref{fig:lazyex}, the final states of the original and variant
      differ, because a written value (in black or white) depended on the value
      of a high confidentiality location. So the resulting trace fails Trace
      Confidentiality. For observable trace confidentiality, we ask whether the
      differing values at address 42 cause any observable distinctions in
      future behavior. To do so, we must undo the variation of formerly high
      confidentiality locations, so that those in yellow match those in red,
      while preserving the updated values where data may have been leaked.
